Addition and removal energies of circular quantum dots

### 2017-04-20 Manuscript/paper.tex
Tweak theme
Add updated rel-slopes figure

In this paper, we analyze the application of several methods to the quantum dot systems, including the Hartree--Fock (HF) method, M\o ller-Plesset perturbation theory (MPPT), the in-medium similarity renormalization group (IM-SRG) method, coupled cluster (CC) theory \cite{PhysRevB.67.045320,heidari:114708,PhysRevB.84.115302}, quasidegenerate perturbation theory (QDPT), and equation-of-motion (EOM) methods.  We compare them against variational and diffusion Monte Carlo (VMC and DMC) \cite{PhysRevB.68.035304,PhysRevB.62.8120,PhysRevB.84.115302,PhysRevB.54.4780} and full configuration interaction (FCI) \cite{olsen2013thesis,JJAP.36.3924,PhysRevB.56.6428,2008arXiv0810.2644K,rontani:124102} results from existing literature.

@@

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-gs2}
  \caption{Ground state energies.}
  \label{fig:gs}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-add2}
  \caption{Addition energies.}
  \label{fig:add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-rm2}
  \caption{Removal energies.  See Figure \ref{fig:add} for details.}
  \label{fig:rm}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-rel-slopes}
  \caption{Comparison of the rate of convergence for various modifications to the Coulomb interaction.  The trends suggest that the short-range part of the interaction has a strong impact on the convergence, whereas the long-range tail has little effect.}
  \label{fig:rel-slopes}
\end{figure}

@@

There are several paths in this diagram through which one can traverse to obtain the output observables.  Among these possibilities, we shall primarily focus on the three combinations: (a) HF + IM-SRG + QDPT, (b) HF + IM-SRG + EOM, and (c) HF + CC + EOM.

@@

In Figure \ref{fig:gs} we display a selection of ground state energies calculated using the three approaches described in Section \ref{subsec:methodology}.  We also include results from DMC \cite{hoegberget2013thesis} and FCI \cite{olsen2013thesis} for comparison.

Both perturbation theory and IM-SRG add significant corrections to the Hartree--Fock results.  Compared with MPPT, IM-SRG generally converges faster, although does not always converge to the DMC result.  Generally speaking, IM-SRG is closer to the DMC result as the number of particles increases.

There are a few cases where the IM-SRG over-corrects the result, leading to a ground state energy that is lower than the exact result.  This is not unexpected given that, unlike Hartree--Fock, IM-SRG is non-variational in the presence of operator truncations, which cause small unitarity violations.  Overcorrection tends to occur when the number of particles is few, or the frequency is very low (high correlation).  However, MPPT does not do much better in the high frequency regimes as it often converges slower than IM-SRG.

>>>>

In this paper, we analyze the application of several methods to the quantum dot systems, including the Hartree--Fock (HF) method, M\o ller-Plesset perturbation theory (MP), the in-medium similarity renormalization group (IM-SRG) method, coupled cluster (CC) theory \cite{PhysRevB.67.045320,heidari:114708,PhysRevB.84.115302}, quasidegenerate perturbation theory (QDPT), and equation-of-motion (EOM) methods.  We compare them against variational and diffusion Monte Carlo (VMC and DMC) \cite{PhysRevB.68.035304,PhysRevB.62.8120,PhysRevB.84.115302,PhysRevB.54.4780} and full configuration interaction (FCI) \cite{olsen2013thesis,JJAP.36.3924,PhysRevB.56.6428,2008arXiv0810.2644K,rontani:124102} results from existing literature.

@@

We will refer to QDPT to third order as ``QDPT3''.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-gs2}
  \caption{Ground state energies for a selection of quantum dot parameters, plotted against the number of shells.}
  \label{fig:gs}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-add2}
  \caption{Addition energies for a selection of quantum dot parameters.  See Figure \ref{fig:gs} for details.}
  \label{fig:add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-rm2}
  \caption{Removal energies for a selection of quantum dot parameters.  See Figure \ref{fig:add} for details.}
  \label{fig:rm}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig-rel-slopes2}
  \caption{The impact of the interaction on convergence of addition and removal energies using IM-SRG + QDPT3.  For clarity, the plot does not distinguish between addition and removal energies.  The horizontal axis shows quantum system parameters, where $N$ is the number of particles and $\omega$ is the oscillator frequency.  The vertical axis shows $|\rho_{15}|$ (``relative slope''), which estimates the rate of convergence at 15 total shells.  The lower the value of $|\rho_{15}|$, the faster the convergence.  The data points are categorized by the interactions.  The trends suggest that the singular short-range part of the interaction has a much stronger impact on the convergence than the long-range tail.}
  \label{fig:rel-slopes}
\end{figure}

@@

There are several paths in this diagram through which one can traverse to obtain the output observables.  Among these possibilities, we shall primarily focus on the three combinations: (a) HF + IM-SRG(2) + QDPT3, (b) HF + IM-SRG(2) + EOM, and (c) HF + CCSD + EOM.

@@

In Figure \ref{fig:gs} we display a selection of ground state energies calculated using the three approaches described in Section \ref{subsec:methodology}.  We also include results from M\o ller--Plesset perturbation theory to second order (MP2), DMC \cite{hoegberget2013thesis}, and FCI \cite{olsen2013thesis} for comparison.

Both perturbation theory and IM-SRG add significant corrections to the Hartree--Fock results.  Compared with MP2, IM-SRG generally converges faster, although does not always converge to the DMC result.  Generally speaking, IM-SRG is closer to the DMC result as the number of particles increases.

There are a few cases where the IM-SRG over-corrects the result, leading to a ground state energy that is lower than the exact result.  This is not unexpected given that, unlike Hartree--Fock, IM-SRG is non-variational in the presence of operator truncations, which cause small unitarity violations.  Overcorrection tends to occur when the number of particles is few, or the frequency is very low (high correlation).  However, MP2 does not do much better in the high frequency regimes as it often converges slower than IM-SRG.
